Emergent dual scaling of riverine biodiversity

### 2021-06-22 manuscript_pnas.Rmd
edits: main manuscript

---
title: "Emergent dual scaling of riverine biodiversity"
output: word_document
csl: pnas.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: references.bib
---

@@

The past two decades have witnessed efforts to illuminate the role of spatial ecosystem structure in shaping biodiversity patterns [@rinaldoRiverNetworksEcological2018; @grantLivingBranchesPopulation2007]. Yet, current schemes have overlooked one of the fundamental rules in branching ecosystems, i.e., the dual scaling law. Here, our unified framework unveiled the surprising consistency of dual scaling across ecological contexts or geographic regions, suggesting the universality of our findings. This discovery brings a paradigm shift in macroecological science.

@@

Under each simulation scenario, we introduced 50 species ($S = 50$) and simulated 1400 time steps of metacommunity dynamics (including 200 time steps for initialization and 200 time steps for burn-in) in `r max(dat_sim$n_rep)` branching networks with the gradients of ecosystem size (the number of habitat patches: `r min(dat_sim$n_patch)` to `r max(dat_sim$n_patch)`) and complexity (branching probability: `r round(min(dat_sim$p_branch), 2)` to `r round(max(dat_sim$p_branch), 2)`). During the first 200 time steps (an initialization period), we initialized the simulation by seeding each habitat patch with populations of each species drawn from a Poisson distribution with a mean of 0.5. We repeated the seeding every 10 time steps during the initialization to allow species to establish their populations if they can grow from low abundances. After the initialization period, we ran 200 time steps as a burn-in period with no seeding to minimize influences of initial conditions. We saved the last 1000 time steps to estimate the temporal averages of $\alpha$, $\beta$, and $\gamma$ diversity. $\beta$ diversity was defined as $\gamma/\alpha$. We removed simulation replicates in which no species established populations over the initial 400 time steps (`r sum(dat_sim$alpha_div == 0)` replicates out of `r nrow(dat_sim)` replicates). Ecosystem size and complexity were drawn randomly from uniform distributions as: $N_p \sim Unif(10, 150)$ and $P_b \sim Unif(0.01, 0.99)$ ($N_p$ was rounded to the nearest integer before running simulations). Values of simulation parameters were summarized in **Table S2**. We performed simulations using R version 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2020].

@@

For the supported model of $\gamma$ diversity, we estimated average predictive comparisons of watershed area and branching probability to compare their effect sizes. The average predictive comparison is an expected change of the response variable per a unit difference of the explanatory variable of interest [@gelmanAveragePredictiveComparisons2007]. Unlike other common metrics of effect size (e.g., standardized regression coefficients), this method provides intuitive yet rigorous estimates of effect size in models with non-linear variable transformations (see **Supporting Information** for further details). We performed statistical analysis using R version 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2020].

>>>>

---
title: "Emergent dual scaling of riverine biodiversity"
output:
  word_document:
    reference_docx: word_format.docx
csl: pnas.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: references.bib
---

@@

The past two decades have witnessed efforts to illuminate the role of spatial ecosystem structure in shaping biodiversity patterns [@rinaldoRiverNetworksEcological2018; @grantLivingBranchesPopulation2007]. Yet, current schemes have overlooked one of the fundamental rules in branching ecosystems, i.e., the dual scaling law. Our unified framework unveiled the surprising consistency of dual scaling across ecological contexts or geographic regions, suggesting the universality of our findings. This discovery, therefore, should bring a paradigm shift in macroecological science.

@@

Under each simulation scenario, we introduced 50 species ($S = 50$) and simulated 1400 time steps of metacommunity dynamics (including 200 time steps for initialization and 200 time steps for burn-in) in `r max(dat_sim$n_rep)` branching networks with the gradients of ecosystem size (the number of habitat patches: `r min(dat_sim$n_patch)` to `r max(dat_sim$n_patch)`) and complexity (branching probability: `r round(min(dat_sim$p_branch), 2)` to `r round(max(dat_sim$p_branch), 2)`). During the first 200 time steps (an initialization period), we initialized the simulation by seeding each habitat patch with populations of each species drawn from a Poisson distribution with a mean of 0.5. We repeated the seeding every 10 time steps during the initialization to allow species to establish their populations if they can grow from low abundances. After the initialization period, we ran 200 time steps as a burn-in period with no seeding to minimize influences of initial conditions. We saved the last 1000 time steps to estimate the temporal averages of $\alpha$, $\beta$, and $\gamma$ diversity. $\beta$ diversity was defined as $\gamma/\alpha$. We removed simulation replicates in which no species established populations over the initial 400 time steps (`r sum(dat_sim$alpha_div == 0)` replicates out of `r nrow(dat_sim)` replicates). Ecosystem size and complexity were drawn randomly from uniform distributions as: $N_p \sim Unif(10, 150)$ and $P_b \sim Unif(0.01, 0.99)$ ($N_p$ was rounded to the nearest integer before running simulations). Values of simulation parameters were summarized in **Table S2**. We performed simulations using R version 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021].

@@

For the supported model of $\gamma$ diversity, we estimated average predictive comparisons of watershed area and branching probability to compare their effect sizes. The average predictive comparison is an expected change of the response variable per a unit difference of the explanatory variable of interest [@gelmanAveragePredictiveComparisons2007]. Unlike other common metrics of effect size (e.g., standardized regression coefficients), this method provides intuitive yet rigorous estimates of effect size in models with non-linear variable transformations (see **Supporting Information** for further details). We performed statistical analysis using R version 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021].
